---
title: "household_data"
output: html_document
date: "2026-04-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r}
# Load the package.
library(datasets)
library(tidyr)
library(stringr)
library(dplyr)
library(rvest)
library(stringr)
library(lubridate)
library(tidytext)
library(ggplot2)
library(knitr)
library(kableExtra)
library(ggthemes)
```

## Themes

```{r}
theme_custom <- function() {
  theme_classic() %+replace%    # Use theme_minimal as the base template
    theme(
      panel.grid.major = element_line(color = 'lightgray', linetype = "dotted")
    )
}
```

## Data Cleaning

### Census

```{r}
census <- read.csv('data/census/census_96_21.csv')

census$QRTYP |> unique()

census |>
  mutate(
    period = year |> factor(levels=year|>unique()),
    LQTYPEDesc = ifelse(QRTYP==9, 4, QRTYP) |> 
      factor(levels=0:4,labels=c('All housing','Public housing','Subsidised housing','Private housing','Others'))
  ) |>
  group_by(period,HHNO) |>
  summarise(
    inc_grp = first(DJHHINC),
    LQTYPEDesc = first(LQTYPEDesc)
  ) |>
  group_by(period,LQTYPEDesc) |>
  summarise(
    inc_grp = median(inc_grp)
  ) |>
  arrange(period,LQTYPEDesc)
```

### Consumption share

```{r}
con_shr <- read.csv('data/household/con_shr.csv')
con_shr <- con_shr |>
  filter(Country.Name %in% c('Hong Kong SAR, China','World','East Asia & Pacific','High income','Middle income','Low income')) |>
  pivot_longer(-c(Country.Name,Country.Code,Indicator.Name,Indicator.Code),names_to = 'period',values_to = 'share') |>
  select(period,Country.Name,share) |>
  mutate(period = str_replace(period,'X','') |> as.numeric())
con_shr <- con_shr |> filter(!is.na(share))
colnames(con_shr) <- c('period','country','share')
con_shr$country <- con_shr$country |> factor(
  levels=c('Hong Kong SAR, China','East Asia & Pacific','World','High income country','Middle income country','Low income country'),
  labels=c('Hong Kong','East Asia & Pacific','World','High income','Middle income','Low income')
  )

head(con_shr)
```

### Consumption

```{r}
cons <- read.csv('data/household/household_expenditure.csv')
cons$GROUPDesc <- cons$GROUPDesc |> 
  factor(levels=
           c('Alcoholic drinks and tobacco','Electricity, gas and water','Clothing and footwear','Miscellaneous goods','Durable goods','Transport','Miscellaneous services','Food','Housing')
         )
cons$LQTYPEDesc <- ifelse(cons$LQTYPEDesc == 'Total', 'All housing', cons$LQTYPEDesc) |>
  factor(levels=c('All housing','Public housing','Subsidised housing','Private housing'))
#cons$period <- cons$period |> factor(labels=cons$period |> unique())
cons <- cons |> select(period,GROUPDesc,LQTYPEDesc,figure)
colnames(cons) <- c('period','goods','housing','expense')

cons |> head()
```

```{r}
cons_gran <- read.csv('data/household/household_expenditure_granular.csv')
cons_gran |> mutate(grp_num = str_split_1(GROUPDesc,'.'))



cons_gran$GROUPDesc <- cons_gran$GROUPDesc |> 
  factor(levels=
           c('Alcoholic drinks and tobacco','Electricity, gas and water','Clothing and footwear','Miscellaneous goods','Durable goods','Transport','Miscellaneous services','Food','Housing')
         )

cons_gran$LQTYPEDesc <- ifelse(cons_gran$LQTYPEDesc == 'Total', 'All housing', cons_gran$LQTYPEDesc) |>
  factor(levels=c('All housing','Public housing','Subsidised housing','Private housing'))
#cons_gran$period <- cons_gran$period |> factor(labels=cons_gran$period |> unique())
cons_gran <- cons_gran |> select(period,GROUPDesc,LQTYPEDesc,figure)
colnames(cons_gran) <- c('period','goods','housing','expense')

cons_gran |> head()
```

### Consumer prices

```{r}
cpi <- read.csv('data/household/consumer_price.csv')
cpi$GROUPDesc <- ifelse(cpi$GROUPDesc == 'Total','All goods',cpi$GROUPDesc)
cpi$GROUPDesc <- cpi$GROUPDesc |> 
  factor(levels=
           c('All goods','Alcoholic drinks and tobacco','Electricity, gas and water','Clothing and footwear','Miscellaneous goods','Durable goods','Transport','Miscellaneous services','Food','Housing')
         )
#cpi$period <- cpi$period |> factor(labels=cpi$period |> unique())
cpi$sv <- cpi$sv |> 
  factor(
    levels=c('CC_CM_1920','A_CM_1920','B_CM_1920','C_CM_1920'),
    labels=c('ccpi','cpi_a','cpi_b','cpi_c')
  )
cpi <- cpi |> pivot_wider(names_from = sv, values_from = figure)
cpi <- cpi |> select(period,GROUPDesc,ccpi,cpi_a,cpi_b,cpi_c)
cpi <- cpi |>
  arrange(period,GROUPDesc) |>
  group_by(GROUPDesc) |>
  mutate(
    ccpi = ccpi / first(ccpi) * 100,
    cpi_a = cpi_a / first(cpi_a) * 100,
    cpi_b = cpi_b / first(cpi_b) * 100,
    cpi_c = cpi_c / first(cpi_c) * 100
  )
colnames(cpi) <- c('period','goods','ccpi','cpi_a','cpi_b','cpi_c')

head(cpi)
```

### Household dist

```{r}
dist <- read.csv('data/household/household_types.csv')
dist$LQTYPEDesc <- dist$LQTYPEDesc |>
  factor(
    levels=c('Total','Public rental housing','Subsidised home ownership housing','Private permanent housing','Temporary housing'),
    labels=c('All housing','Public housing','Subsidised housing','Private housing','Temporary housing')
    )
#dist$period <- dist$period |> factor(labels=dist$period |> unique())
dist <- dist |> select(period,LQTYPEDesc,svDesc,figure)
colnames(dist) <- c('period','LQTYPEDesc','svDesc','number')

head(dist)
```

### Household size

```{r}
size <- read.csv('data/household/household_sizes.csv')
size$LQTYPEDesc <- size$LQTYPEDesc |>
  factor(
    levels=c('Total','Public rental housing','Subsidised home ownership housing','Private permanent housing','Temporary housing'),
    labels=c('All housing','Public housing','Subsidised housing','Private housing','Temporary housing')
    )
#size$period <- size$period |> factor(labels=size$period |> unique())
size <- size |> select(period,LQTYPEDesc,figure)
colnames(size) <- c('period','LQTYPEDesc','person')

head(size)
```

### Household income

```{r}

income <- read.csv('data/household/household_income.csv')
income$LQTYPEDesc <- income$LQTYPEDesc |>
  factor(
    levels=c('Total','Public rental housing','Subsidised home ownership housing','Private permanent housing','Temporary housing'),
    labels=c('All housing','Public housing','Subsidised housing','Private housing','Temporary housing')
    )
#income$period <- income$period |> factor(labels=income$period |> unique())
income$sv <- income$sv |> 
  factor(
    levels=c('MED_DH_INC','MED_DH_INC_XEI'),
    labels=c('inc','inc_eco')
  )
income <- income |> pivot_wider(names_from = sv, values_from = figure)
income <- income |> select(period,LQTYPEDesc,inc,inc_eco)
colnames(income) <- c('period','housing','med_inc','med_inc_eco')


agg_inc <- read.csv('data/household/income.csv')
colnames(agg_inc) <- c('period','housing','med_inc','med_inc_eco')
agg_inc$housing <-'All housing' |> as.factor()

income <- rbind(income,agg_inc) |> arrange(period,housing)

head(income)
```

## Aggregate

```{r}
con_shr |>
  filter(between(period,1999,2019)) |>
  filter(country %in% c('Hong Kong','High income','Middle income','Low income')) |>
  ggplot() +
  geom_line(aes(x=period,y=share,color=country),linewidth=1) +
  theme_custom()+
  labs(
    title = "Private consumption share of GDP",
    subtitle = "(by income group, 1999 - 2019)",
    caption = "Data from World Bank",
    x = "Year",
    y = "%",
    color = 'Income group'
  )

con_shr |>
  filter(between(period,1999,2019)) |>
  filter(country %in% c('Hong Kong','World','East Asia & Pacific')) |>
  ggplot() +
  geom_line(aes(x=period,y=share,color=country),linewidth=1) +
  theme_custom()+
  labs(
    title = "Private consumption share of GDP",
    subtitle = "(by location, 1999 - 2019)",
    caption = "Data from World Bank",
    x = "Year",
    y = "%",
    color = 'Location'
  )

```

### Cross country share

```{r}

```

## All households

### Income vs expenditure

```{r}
exp <- cons |> 
  group_by(period,housing) |>
  summarise(expense = sum(expense))

inc_exp <- income |> merge(exp, by=c('period','housing'))
```

```{r}
inc_exp |>
  filter(between(period,1999,2019)) |>
  filter(housing == 'All housing') |>
  pivot_longer(c(med_inc,med_inc_eco,expense),names_to='typ',values_to = 'val') |>
  filter(typ!='med_inc' & !is.na(val)) |>
  mutate(typ=case_when(typ=='expense'~'Average expenditure',typ=='med_inc'~'Median income', typ=='med_inc_eco'~'Median income')) |>
  ggplot() +
  #geom_bar(aes(x=period,y=pct_chg,fill=typ),color='white',stat='identity',position='dodge') +
  geom_line(aes(x=period,y=val,color=typ),linewidth = 1) +
  theme_custom() +
  labs(
    title = "Household income vs expenditure",
    subtitle = "(1999 - 2019)",
    caption = "Data from C&SD",
    x = "Year",
    y = "HK$",
    color = ''
  )
```

### Expenditure trends

```{r}
cons |>
  filter(housing == 'All housing') |>
  arrange(goods,period) |>
  ggplot() +
  geom_bar(aes(x=period,y=expense,fill=goods),color='white',stat='identity',position='stack') +
  theme_custom() +
  labs(
    title = "Household expenditure",
    subtitle = "(1999 - 2019)",
    caption = "Data from C&SD",
    x = "Year",
    y = "HK$",
    fill = 'Goods'
  )

cons |>
  filter(housing == 'All housing') |>
  arrange(goods,period) |>
  group_by(period) |>
  mutate(tot_exp = sum(expense)) |>
  group_by(goods) |>
  summarise(
    pct_change = (last(expense) - first(expense)) / (first(tot_exp)) * 100,
    ) |>
  ggplot() +
  geom_bar(aes(x=reorder(goods,pct_change),y=pct_change,fill=goods),color='white',stat='identity',show.legend=FALSE) +
  geom_text(
    aes(x = reorder(goods, pct_change), y = pct_change, 
        label = ifelse(pct_change > 1, sprintf("%+.0f%%", pct_change), '')
        ),
    hjust = 1.1,  # Adjusts distance from the end of the bar
    size = 3,
    color ='white',
    fontface = 'bold'
  ) +
  coord_flip() + 
  theme_custom() +
  theme(
    axis.text.y = element_text(angle = 30, vjust = 0, hjust = 1)
  ) +
  labs(
    title = "Household expenditure, total % change",
    subtitle = "(1999 - 2019)",
    caption = "Data from C&SD",
    x = "Goods",
    y = "%Δ",
  )

cons |>
  filter(housing == 'All housing') |>
  arrange(period,goods) |>
  pivot_wider(goods,names_from = period,values_from = expense) |>
  kable(
    caption = "Household expenditure",
    digits = 2
  )

cons |>
  filter(housing == 'All housing') |>
  arrange(goods,period) |>
  group_by(period) |>
  mutate(tot_exp = sum(expense)) |>
  group_by(goods) |>
  summarise(
    pct_change = (last(expense) - first(expense)) / (first(tot_exp)) * 100,
    ) |>
  kable(
    caption = "Household expenditure",
    digits = 2
  )
```

### Composition of expenditure

```{r}
cons |>
  filter(housing == 'All housing') |>
  group_by(period) |>
  mutate(share = expense / sum(expense) * 100) |>
  arrange(goods,period) |>
  ggplot() +
  geom_area(aes(x=period,y=share,fill=goods),color='white',stat='identity',position='stack') +
  theme_custom() +
  labs(
    title = "Composition of household expenditure",
    subtitle = "(1999 - 2019)",
    caption = "Data from C&SD",
    x = "Year",
    y = "%",
    fill = 'Goods'
  )

cons |>
  filter(housing == 'All housing') |>
  group_by(period) |>
  mutate(share = expense / sum(expense) * 100) |>
  arrange(goods,period) |>
  group_by(goods) |>
  summarise(change = last(share) - first(share)) |>
  ggplot() +
  geom_bar(aes(x=reorder(goods,change),y=change,fill=goods),stat='identity',color='white',show.legend=FALSE) +
  geom_text(
    aes(x = reorder(goods, change), y = change, 
        label = ifelse(change > 1, sprintf("%+.0f%%", change), '')
        ),
    hjust = 1.1,  # Adjusts distance from the end of the bar
    size = 3,
    color ='white',
    fontface = 'bold'
  ) +
  geom_text(
    aes(x = reorder(goods, change), y = change, 
        label = ifelse(change < -1, sprintf("%+.0f%%", change), '')
        ),
    hjust = -0.2,  # Adjusts distance from the end of the bar
    size = 3,
    color ='white',
    fontface = 'bold'
  ) +
  coord_flip() + 
  theme_custom() +
  theme(
    axis.text.y = element_text(angle = 30, vjust = 0, hjust = 1)
  ) +
  labs(
    title = "Composition of household expenditure, total change",
    subtitle = "(1999 - 2019)",
    caption = "Data from C&SD",
    x = "Goods",
    y = "Δ%",
  )

cons |>
  filter(housing == 'All housing') |>
  group_by(period) |>
  mutate(share = expense / sum(expense) * 100) |>
  arrange(goods,period) |>
  pivot_wider(goods,names_from = period,values_from = share) |>
  kable(
    caption = "Composition of household expenditure",
    digits = 2
  )

cons |>
  filter(housing == 'All housing') |>
  group_by(period) |>
  mutate(share = expense / sum(expense) * 100) |>
  arrange(goods,period) |>
  group_by(goods) |>
  summarise(change = last(share) - first(share)) |>
  kable(
    caption = "Composition of household expenditure",
    digits = 2
  )
```

### Nominal vs real

```{r}
cons_cpi <- cons |>
  merge(cpi,by=c('period','goods'))
cons_cpi
```

```{r}
cons_cpi |>
  filter(between(period,1999,2019)) |>
  filter(housing == 'All housing') |>
  arrange(goods,period) |>
  ggplot(aes(x=period)) +
  geom_bar(aes(y=expense,fill=goods),color='white',stat='identity',position='stack') +
  geom_line(
    data = . %>% group_by(period) %>% summarise(ccpi = mean(ccpi)),
    aes(y=ccpi * 260),
    linetype = 'dashed',
    linewidth = 1,
  ) +
  geom_point(
    data = . %>% group_by(period) %>% summarise(ccpi = mean(ccpi)),
    aes(y=ccpi * 260),
    shape = 4,
    size = 3,
    stroke = 1,
  ) +
  scale_y_continuous(
    name = "Nominal expenditure",
    sec.axis = sec_axis(~ . / 260, name = "Composite CPI (dashed)") # Divide to restore labels
  ) +
  theme_custom() +
  labs(
    title = "Household expenditure vs consumer price",
    subtitle = "(1999 - 2019)",
    caption = "Data from C&SD",
    x = "Year",
    fill = 'Goods'
  )

cons_cpi |>
  filter(between(period,1999,2019)) |>
  filter(housing=='All housing') |>
  arrange(goods,period) |>
  group_by(goods) |>
  summarise(change = last(ccpi) - first(ccpi)) |>
  ggplot() +
  geom_bar(aes(x=reorder(goods,change),y=change,fill=goods),stat='identity',color='white',show.legend=FALSE) +
  coord_flip() +
  geom_text(
    aes(x = reorder(goods, change), y = change, 
        label = ifelse(change > 10, sprintf("%+.0f", change), '')
        ),
    hjust = 1.1, size = 3, color ='white', fontface = 'bold'
  ) +
  geom_text(
    aes(x = reorder(goods, change), y = change, 
        label = ifelse(change < -10, sprintf("%+.0f", change), '')
        ),
    hjust = -0.2, size = 3, color ='white', fontface = 'bold'
  ) +
  theme_custom() +
  theme(
    axis.text.y = element_text(angle = 30, vjust = 0, hjust = 1)
  ) +
  labs(
    title = "Consumer prices (1999 = 100), total change",
    subtitle = "(1999 - 2019)",
    caption = "Data from C&SD",
    x = "Goods",
    y = "ΔCPI"
  )

cons_cpi |>
  filter(between(period,1999,2019)) |>
  filter(housing=='All housing') |>
  mutate(real = expense / ccpi) |>
  arrange(goods,period) |>
  group_by(period) |>
  mutate(tot_real = sum(real)) |>
  group_by(goods) |>
  summarise(
    pct_change = (last(real) - first(real)) / (first(tot_real)) * 100,
    ) |>
  ggplot()+
  geom_bar(aes(x=reorder(goods,pct_change),y=pct_change,fill=goods),color='white',stat='identity',show.legend=FALSE) +
  geom_text(
    aes(x = reorder(goods, pct_change), y = pct_change, 
        label = ifelse(pct_change > 1, sprintf("%+.0f%%", pct_change), '')
        ),
    hjust = 1.1, size = 3, color ='white', fontface = 'bold'
  ) +
  geom_text(
    aes(x = reorder(goods, pct_change), y = pct_change, 
        label = ifelse(pct_change < -1, sprintf("%+.0f%%", pct_change), '')
        ),
    hjust = -0.2, size = 3, color ='white', fontface = 'bold'
  ) +
  coord_flip() + 
  theme_custom() +
  theme(
    axis.text.y = element_text(angle = 30, vjust = 0, hjust = 1)
  ) +
  labs(
    title = "Real expenditure, total % change",
    subtitle = "(1999 - 2019)",
    caption = "Data from C&SD",
    x = "Goods",
    y = "%Δ",
  )

cons_cpi |>
  filter(between(period,1999,2019)) |>
  filter(housing=='All housing') |>
  mutate(real = expense / ccpi) |>
  arrange(goods,period) |>
  group_by(period) |>
  mutate(tot_real = sum(real)) |>
  group_by(goods) |>
  summarise(
    pct_change = (last(real) - first(real)) / (first(tot_real)) * 100,
    ) |>
  kable(
    caption = "Real household expenditure",
    digits = 2
  )
```

## By housing type

```{r}
inc_sampled <- income |>
  filter(period %in% cons$period)

cons |> merge(inc_sampled, by=c('period','LQTYPEDesc'))
```

### Income vs expenditure

```{r}
inc_exp |>
  filter(between(period,1999,2019)) |>
  filter(housing != 'All housing') |>
  pivot_longer(c(med_inc,med_inc_eco,expense),names_to='typ',values_to = 'val') |>
  filter(typ!='med_inc' & !is.na(val)) |>
  mutate(typ=case_when(typ=='expense'~'Average expenditure',typ=='med_inc'~'Median income', typ=='med_inc_eco'~'Median income')) |>
  ggplot() +
  #geom_bar(aes(x=period,y=pct_chg,fill=typ),color='white',stat='identity',position='dodge') +
  geom_line(aes(x=period,y=val,color=typ),linewidth = 1) +
  facet_wrap(~housing, nrow=1) +
  theme_custom() +
  labs(
    title = "Household income vs expenditure",
    subtitle = "(by housing type, 1999 - 2019)",
    caption = "Data from C&SD",
    x = "Year",
    y = "HK$",
    color = ''
  )
```

### Expenditure trends

```{r}
cons |>
  filter(housing != 'All housing') |>
  arrange(goods,period) |>
  ggplot() +
  geom_bar(aes(x=period,y=expense,fill=goods),color='white',stat='identity',position='stack') +
  facet_wrap(~housing,nrow=1) +
  theme_custom() +
  theme(
    axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1)
  ) +
  labs(
    title = "Household expenditure",
    subtitle = "(by housing type, 1999 - 2019)",
    caption = "Data from C&SD",
    x = "Year",
    y = "HK$",
    fill = 'Goods'
  )

cons |>
  filter(housing != 'All housing') |>
  arrange(goods,period) |>
  group_by(period,housing) |>
  mutate(tot_exp = sum(expense)) |>
  group_by(goods,housing) |>
  summarise(
    pct_change = (last(expense) - first(expense)) / (first(tot_exp)) * 100,
    ) |>
  ggplot() +
  geom_bar(aes(x=reorder(goods,pct_change),y=pct_change,fill=goods),color='white',stat='identity',show.legend=FALSE) +
  facet_wrap(~housing,nrow=1) +
  geom_text(
    aes(x = reorder(goods, pct_change), y = pct_change, 
        label = ifelse(pct_change > 4, sprintf("%+.0f%%", pct_change), '')
        ),
    hjust = 1.1,  # Adjusts distance from the end of the bar
    size = 3,
    color ='white',
    fontface = 'bold'
  ) +
  coord_flip() + 
  theme_custom() +
  theme(
    axis.text.y = element_text(angle = 30, vjust = 0, hjust = 1)
  ) +
  labs(
    title = "Household expenditure, total % change",
    subtitle = "(by housing type, 1999 - 2019)",
    caption = "Data from C&SD",
    x = "Goods",
    y = "%Δ",
  )

cons |>
  filter(housing != 'All housing') |>
  arrange(period,housing,goods) |>
  pivot_wider(goods,names_from = period,values_from = expense) |>
  kable(
    caption = "Household expenditure",
    digits = 2
  )

cons |>
  filter(housing != 'All housing') |>
  arrange(goods,period) |>
  group_by(period,housing) |>
  mutate(tot_exp = sum(expense)) |>
  group_by(goods,housing) |>
  summarise(
    pct_change = (last(expense) - first(expense)) / (first(tot_exp)) * 100,
    ) |>
  kable(
    caption = "Household expenditure",
    digits = 2
  )
```

### Composition of expenditure

```{r}
cons |>
  filter(housing != 'All housing') |>
  group_by(period,housing) |>
  mutate(share = expense / sum(expense) * 100) |>
  arrange(goods,period) |>
  ggplot() +
  geom_area(aes(x=period,y=share,fill=goods),color='white',stat='identity',position='stack') +
  facet_wrap(~housing) +
  theme_custom() +
  theme(
    axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1)
  ) +
  labs(
    title = "Composition of household expenditure",
    subtitle = "(by housing type, 1999 - 2019)",
    caption = "Data from C&SD",
    x = "Year",
    y = "%",
    fill = 'Goods'
  )

cons |>
  filter(housing != 'All housing') |>
  group_by(period,housing) |>
  mutate(share = expense / sum(expense) * 100) |>
  arrange(goods,housing,period) |>
  group_by(goods,housing) |>
  summarise(change = last(share) - first(share)) |>
  ggplot() +
  geom_bar(aes(x=reorder(goods,change),y=change,fill=goods),stat='identity',color='white',show.legend=FALSE) +
  geom_text(
    aes(x = reorder(goods, change), y = change, 
        label = ifelse(change > 1.5, sprintf("%+.0f%%", change), '')
        ),
    hjust = 1.1,  # Adjusts distance from the end of the bar
    size = 3,
    color ='white',
    fontface = 'bold'
  ) +
  geom_text(
    aes(x = reorder(goods, change), y = change, 
        label = ifelse(change < -1.5, sprintf("%+.0f%%", change), '')
        ),
    hjust = -0.2,  # Adjusts distance from the end of the bar
    size = 3,
    color ='white',
    fontface = 'bold'
  ) +
  coord_flip() + 
  facet_wrap(~housing) +
  theme_custom() +
  theme(
    axis.text.y = element_text(angle = 30, vjust = 0, hjust = 1)
  ) +
  labs(
    title = "Composition of household expenditure, total change",
    subtitle = "(by housing type, 1999 - 2019)",
    caption = "Data from C&SD",
    x = "Goods",
    y = "Δ%",
  )

cons |>
  filter(housing != 'All housing') |>
  group_by(period,housing) |>
  mutate(share = expense / sum(expense) * 100) |>
  arrange(goods,period) |>
  pivot_wider(goods,housing,names_from = period,values_from = share) |>
  kable(
    caption = "Composition of household expenditure",
    digits = 2
  )

cons |>
  filter(housing != 'All housing') |>
  group_by(period,housing) |>
  mutate(share = expense / sum(expense) * 100) |>
  arrange(goods,period) |>
  group_by(goods,housing) |>
  summarise(change = last(share) - first(share)) |>
  kable(
    caption = "Composition of household expenditure",
    digits = 2
  )
```

### Nominal vs real

```{r}
cons_cpi |>
  filter(between(period,1999,2019)) |>
  filter(housing!='All housing') |>
  mutate(real = expense / ccpi) |>
  arrange(goods,period) |>
  group_by(period,housing) |>
  mutate(tot_real = sum(real)) |>
  group_by(goods,housing) |>
  summarise(
    pct_change = (last(real) - first(real)) / (first(tot_real)) * 100,
    ) |>
  ggplot()+
  geom_bar(aes(x=reorder(goods,pct_change),y=pct_change,fill=goods),color='white',stat='identity',show.legend=FALSE) +
  geom_text(
    aes(x = reorder(goods, pct_change), y = pct_change, 
        label = ifelse(pct_change > 3, sprintf("%+.0f%%", pct_change), '')
        ),
    hjust = 1.1, size = 3, color ='white', fontface = 'bold'
  ) +
  geom_text(
    aes(x = reorder(goods, pct_change), y = pct_change, 
        label = ifelse(pct_change < -3, sprintf("%+.0f%%", pct_change), '')
        ),
    hjust = -0.2, size = 3, color ='white', fontface = 'bold'
  ) +
  coord_flip() + 
  facet_wrap(~housing,nrow=1)+
  theme_custom() +
  theme(
    axis.text.y = element_text(angle = 30, vjust = 0, hjust = 1)
  ) +
  labs(
    title = "Real expenditure, total % change",
    subtitle = "(by housing type, 1999 - 2019)",
    caption = "Data from C&SD",
    x = "Goods",
    y = "%Δ",
  )
```

## Cross-country

```{r}

```
