---
title: "job_data"
output: html_document
date: "2026-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r}
# Load the package.
library(datasets)
library(dslabs)
library(tidyr)
library(stringr)
library(dplyr)
library(rvest)
library(stringr)
library(lubridate)
library(janeaustenr)
library(tidytext)
library(textdata)
library(ggplot2)
library(latex2exp)
library(knitr)
library(zoo)
```

## Data Cleaning

```{r}
df <- read.csv('data/job.csv',sep = ',',na.strings = c('N.A.','[-§32]','[*1]'))
head(df)

colnames(df) <- paste0(df[2,],' - ',df[3,]) |> str_replace('^\\s{2}-\\s$','') |> paste0(df[4,])
df |> str()

df <- df[5:2270,-c(5,7,9)]
df <- df |> mutate(
  ind_des = str_trunc(`Industry section`,70) |> as.factor(),
  ind = str_replace(`Industry section`,': .*','') |> as.factor()
  )
df$Year <- df$Year |> as.numeric()
df[2:3] <- df[2:3] |> lapply(as.factor)
df[4:7] <- df[4:7] |> lapply(as.numeric)
head(df)

df <- df |>
  mutate(
    Date = mdy(paste0(Month,'1, ',Year)),
    `Vacancy rate (1) - (%)` = round(`Number of vacancies - No.`/(`Number of vacancies - No.`+`Number of persons engaged - No.`)*100,2)
  )
df <- df |>
  arrange(Date)

df |> str()

# All industries
all_ind <- df |>
  filter(`Industry section` == 'All industry sections covered')
# Grouped industries
agg_ind <- df |>
  filter(
    `Industry section` %in% 
      c('B, D & E: Mining and quarrying; and electricity and gas supply, and waste management',
        'G: Import/export, wholesale and retail trades',
        'M & N: Professional and business services',
        'P - S: Social and personal services'
        )
  )
# Granular industries without groups
gran_ind <- df |>
  filter(
    `Industry section` != 'All industry sections covered' &
    ! `Industry section` %in% 
      c('B: Mining and quarrying',
        'D & E: Electricity and gas supply, and waste management',
        #'B, D & E: Mining and quarrying; and electricity and gas supply, and waste management',
        'G: Import/export, wholesale and retail trades',
        'M & N: Professional and business services',
        'P - S: Social and personal services'
        )
  )
# Official groupings
agg_ind <- df |>
  filter(
    `Industry section` %in% 
      c('C: Manufacturing',
        'B, D & E: Mining and quarrying; and electricity and gas supply, and waste management',
        'F: Construction sites (manual workers only)',
        'G: Import/export, wholesale and retail trades',
        'H & J: Transportation, storage, postal and courier services; and information and communications',
        'H: Transportation, storage, postal and courier services',
        'J: Information and communicationsJ: Information and communications',
        'I: Accommodation and food services',
        'K & L: Financing and insurance, and real estate',
        'K: Financing and insurance',
        'L: Real estate',
        'M & N: Professional and business services',
        'P - S: Social and personal services'
        )
  )
```

## Summary of labour market

```{r}
# Summary of labour market
sum_lab <- gran_ind |>
  group_by(Date) |>
  summarise(
    tot_est = sum(`Number of establishments - No.`),
    tot_eng = sum(`Number of persons engaged - No.`),
    tot_vac = sum(`Number of vacancies - No.`)
  ) |>
  mutate(
    tot_lab = tot_eng + tot_vac,
    lab_per_est = tot_lab / tot_est,
    eng_per_est = tot_eng / tot_est,
    vac_per_est = tot_vac / tot_est,
    vac_rate = tot_vac / tot_lab * 100
  )

sum_lab_long <- sum_lab |>
  pivot_longer(-Date)

head(sum_lab)
```

### Establishment

```{r}
# Total establishment trend
sum_lab_long |>
  filter(name == 'tot_est') |>
  ggplot(aes(x=Date,y=value)) +
  geom_line() +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Number of establishments in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of establisments"
  ) +
  theme_classic()

# Total establishment changes
sum_lab_long |>
  filter(name == 'tot_est') |>
  mutate(
    value = value - lag(value)
  ) |>
  mutate(
    updown = ifelse(value >= 0, 'Increase','Decrease')
  ) |>
  drop_na() |>
  ggplot(aes(x=Date,y=value,fill=updown)) +
  scale_fill_manual(values = c("Increase" = "chartreuse", "Decrease" = "brown1"), guide = 'none') +
  geom_bar(stat = 'identity') +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Change in number of establishments in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of establisments"
  ) +
  theme_classic()
```

### Engaged persons

```{r}
# Number of engagement
sum_lab_long |>
  filter(name == 'tot_eng') |>
  ggplot(aes(x=Date,y=value)) +
  geom_line() +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Number of engaged persons in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()

# Number of engagement changes
sum_lab_long |>
  filter(name == 'tot_eng') |>
  mutate(
    value = value - lag(value)
  ) |>
  mutate(
    updown = ifelse(value >= 0, 'Increase','Decrease')
  ) |>
  drop_na() |>
  ggplot(aes(x=Date,y=value,fill=updown)) +
  scale_fill_manual(values = c("Increase" = "chartreuse", "Decrease" = "brown1"), guide = 'none') +
  geom_bar(stat = 'identity') +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Number of engaged persons in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()

```

### Vacancies

```{r}
# Number of vacancies
sum_lab_long |>
  filter(name == 'tot_vac') |>
  ggplot(aes(x=Date,y=value)) +
  geom_line() +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Number of vacancies in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()

# Number of vacancies changes
sum_lab_long |>
  filter(name == 'tot_vac') |>
  mutate(
    value = value - lag(value)
  ) |>
  mutate(
    updown = ifelse(value >= 0, 'Increase','Decrease')
  ) |>
  drop_na() |>
  ggplot(aes(x=Date,y=value,fill=updown)) +
  scale_fill_manual(values = c("Increase" = "chartreuse", "Decrease" = "brown1"), guide = 'none') +
  geom_bar(stat = 'identity') +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Number of vacancies in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()
```

### Correlation

```{r}
# Establishment vs engagement
sum_lab |>
  select(Date,tot_est,tot_eng) |>
  ggplot(aes(x = Date)) +
  geom_line(aes(y = tot_est, color = "LHS")) +
  geom_line(aes(y = tot_eng / 7.5, color = "RHS")) + 
  scale_y_continuous(
    name = "Establishments",
    sec.axis = sec_axis(~ . * 7.5, name = 'Engaged persons')
  ) +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Number of establishments vs engaged persons in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
  ) +
  theme_classic()

# Establishment vs vacancies
sum_lab |>
  select(Date,tot_est,tot_vac) |>
  ggplot(aes(x = Date)) +
  geom_line(aes(y = tot_est, color = "LHS")) +
  geom_line(aes(y = log(tot_vac) * 30000, color = "RHS")) + 
  scale_y_continuous(
    name = "Establishments",
    sec.axis = sec_axis(~ . / 30000, name = 'log(Vacancies)')
  ) +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Number of establishments vs vacancies in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
  ) +
  theme_classic()

# Engagement vs vacancies
sum_lab |>
  select(Date,tot_eng,tot_vac) |>
  ggplot(aes(x = Date)) +
  geom_line(aes(y = tot_eng, color = "LHS")) +
  geom_line(aes(y = log(tot_vac) * 230000, color = "RHS")) + 
  scale_y_continuous(
    name = "Engaged persons",
    sec.axis = sec_axis(~ . / 230000, name = 'log(Vacancies)')
  ) +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Number of engaged persons vs vacancies in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
  ) +
  theme_classic()
```

### Headcounts per establishment

```{r}
# Headcounts per establishment
sum_lab_long |>
  filter(name == 'lab_per_est') |>
  ggplot(aes(x=Date,y=value)) +
  geom_line() +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Headcounts per establishment in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()

# Headcounts per establishment changes
sum_lab_long |>
  filter(name == 'lab_per_est') |>
  mutate(
    value = value - lag(value)
  ) |>
  mutate(
    updown = ifelse(value >= 0, 'Increase','Decrease')
  ) |>
  drop_na() |>
  ggplot(aes(x=Date,y=value,fill=updown)) +
  scale_fill_manual(values = c("Increase" = "chartreuse", "Decrease" = "brown1"), guide = 'none') +
  geom_bar(stat = 'identity') +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Change in headcounts per establishments in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()
```

### Engagement per establishment

```{r}
# Engagement per establishment
sum_lab_long |>
  filter(name == 'eng_per_est') |>
  ggplot(aes(x=Date,y=value)) +
  geom_line() +
  labs(
    title = "Engaged persons per establishment in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()

# Engagement per establishment changes
sum_lab_long |>
  filter(name == 'eng_per_est') |>
  mutate(
    value = value - lag(value)
  ) |>
  mutate(
    updown = ifelse(value >= 0, 'Increase','Decrease')
  ) |>
  drop_na() |>
  ggplot(aes(x=Date,y=value,fill=updown)) +
  scale_fill_manual(values = c("Increase" = "chartreuse", "Decrease" = "brown1"), guide = 'none') +
  geom_bar(stat = 'identity') +
  labs(
    title = "Change in engaged person per establishments in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()
```

### Vacancies per establishment

```{r}
# Vacancies per establishment
sum_lab_long |>
  filter(name == 'vac_per_est') |>
  ggplot(aes(x=Date,y=value)) +
  geom_line() +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Vacancies per establishment in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()

# Engagement per establishment changes
sum_lab_long |>
  filter(name == 'vac_per_est') |>
  mutate(
    value = value - lag(value)
  ) |>
  mutate(
    updown = ifelse(value >= 0, 'Increase','Decrease')
  ) |>
  drop_na() |>
  ggplot(aes(x=Date,y=value,fill=updown)) +
  scale_fill_manual(values = c("Increase" = "chartreuse", "Decrease" = "brown1"), guide = 'none') +
  geom_bar(stat = 'identity') +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Change in vacancies per establishments in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons"
  ) +
  theme_classic()
```

### Vacancy rate

```{r}
# Vacancy rate
sum_lab_long |>
  filter(name == 'vac_rate') |>
  ggplot(aes(x=Date,y=value)) +
  geom_line() +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Vacancy rate in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "%"
  ) +
  theme_classic()

# Vacancy rate changes
sum_lab_long |>
  filter(name == 'vac_rate') |>
  mutate(
    value = value - lag(value)
  ) |>
  mutate(
    updown = ifelse(value >= 0, 'Increase','Decrease')
  ) |>
  drop_na() |>
  ggplot(aes(x=Date,y=value,fill=updown)) +
  scale_fill_manual(values = c("Increase" = "chartreuse", "Decrease" = "brown1"), guide = 'none') +
  geom_bar(stat = 'identity') +
  geom_vline(
    xintercept = as.Date("2019-06-01"), 
    linetype = "dashed", 
    color = "red", 
    linewidth = 0.5
    ) +
  labs(
    title = "Change in vacancy rate per establishments in all industries",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "%"
  ) +
  theme_classic()
```

## By industry labour market

```{r}
# By industry labour market
ind_lab <- gran_ind |>
  group_by(Date,ind,ind_des) |>
  summarise(
    tot_est = sum(`Number of establishments - No.`),
    tot_eng = sum(`Number of persons engaged - No.`),
    tot_vac = sum(`Number of vacancies - No.`)
  ) |>
  mutate(
    tot_lab = tot_eng + tot_vac,
    lab_per_est = tot_lab / tot_est,
    eng_per_est = tot_eng / tot_est,
    vac_per_est = tot_vac / tot_est,
    vac_rate = tot_vac / tot_lab * 100
  )

ind_lab_long <- ind_lab |>
  pivot_longer(-c(Date,ind,ind_des))

head(ind_lab)
head(ind_lab_long)
```

```{r}
ind_lab |> group_by(ind) |> summarise(des = first(ind_des))
```

### Establishment

```{r}
# Industry establishment trend
ind_lab_long |>
  filter(name == 'tot_est') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of establishments by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of establisments",
    color = 'Industry'
  ) +
  theme_classic()

# Industry establishment changes
ind_lab_long |>
  filter(name == 'tot_est') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of establishments by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry establishment changes
ind_lab_long |>
  filter(name == 'tot_est') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

### Engaged persons

```{r}
# Industry engagement trend
ind_lab_long |>
  filter(name == 'tot_eng') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of engaged persons by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry engagement changes
ind_lab_long |>
  filter(name == 'tot_eng') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of engaged persons by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry engagement changes
ind_lab_long |>
  filter(name == 'tot_eng') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

### Vacancies

```{r}
# Industry vacancies trend
ind_lab_long |>
  filter(name == 'tot_vac') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of vacancies by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry vacancies changes
ind_lab_long |>
  filter(name == 'tot_vac') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of vacancies by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry vacancies changes
ind_lab_long |>
  filter(name == 'tot_vac') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

### Headcounts per establishment

```{r}

```

```{r}
# Industry headcounts trend
ind_lab_long |>
  filter(name == 'lab_per_est') |>
  filter(!(ind %in% c('B, D & E', 'F'))) |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of headcounts per establishment by industry (ex B,D,E,F)",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry headcounts changes
ind_lab_long |>
  filter(name == 'lab_per_est') |>
  filter(!(ind %in% c('B, D & E', 'F'))) |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of headcounts per establishment by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry headcounts changes
ind_lab_long |>
  filter(name == 'lab_per_est') |>
  filter(!(ind %in% c('B, D & E', 'F'))) |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

### Engaged persons per establishment

```{r}
# Industry engagement per establishment trend
ind_lab_long |>
  filter(name == 'eng_per_est') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of engaged persons per establishment by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry engagement per establishment changes
ind_lab_long |>
  filter(name == 'eng_per_est') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of engaged persons per establishment by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry engagement per establishment changes
ind_lab_long |>
  filter(name == 'eng_per_est') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

### Vacancies per establishment

```{r}
# Industry vacancies per establishment trend
ind_lab_long |>
  filter(name == 'vac_per_est') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of vacancies per establishment by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry vacancies per establishment changes
ind_lab_long |>
  filter(name == 'vac_per_est') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of vacancies per establishment by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry vacancies per establishment changes
ind_lab_long |>
  filter(name == 'vac_per_est') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

### Vacancy rate

```{r}
# Industry vacancies rate trend
ind_lab_long |>
  filter(name == 'vac_rate') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Vacancy rate by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "%",
    color = 'Industry'
  ) +
  theme_classic()

# Industry vacancy rate changes
ind_lab_long |>
  filter(name == 'vac_rate') |>
  group_by(ind) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in vacancy rate by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry vacancy rate changes
ind_lab_long |>
  filter(name == 'vac_rate') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  arrange(-change) |>
  print()
```

## Pre/post-2019

```{r}
pre19 <- ind_lab_long |>
  filter(Date < '2019-01-01')
pos19 <- ind_lab_long |>
  filter(Date >= '2019-01-01')
```

### Establishment

```{r}
# Industry establishment trend
pre19 |>
  filter(name == 'tot_est') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of establishments by industry",
    subtitle = "(2000 - 2019)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of establisments",
    color = 'Industry'
  ) +
  theme_classic()

# Industry establishment changes
pre19 |>
  filter(name == 'tot_est') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of establishments by industry",
    subtitle = "(2000 - 2019)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry establishment changes
pre19 |>
  filter(name == 'tot_est') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

```{r}
# Industry establishment trend
pos19 |>
  filter(name == 'tot_est') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of establishments by industry",
    subtitle = "(2019 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of establisments",
    color = 'Industry'
  ) +
  theme_classic()

# Industry establishment changes
pos19 |>
  filter(name == 'tot_est') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of establishments by industry",
    subtitle = "(2019 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry establishment changes
pos19 |>
  filter(name == 'tot_est') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

### Engaged persons

```{r}
# Industry engagement trend
pre19 |>
  filter(name == 'tot_eng') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of engaged persons by industry",
    subtitle = "(2000 - 2019)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry engagement changes
pre19 |>
  filter(name == 'tot_eng') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of engaged persons by industry",
    subtitle = "(2000 - 2019)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry engagement changes
pre19 |>
  filter(name == 'tot_eng') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

```{r}
# Industry engagement trend
pos19 |>
  filter(name == 'tot_eng') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of engaged persons by industry",
    subtitle = "(2019 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry engagement changes
pos19 |>
  filter(name == 'tot_eng') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of engaged persons by industry",
    subtitle = "(2019 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry engagement changes
pos19 |>
  filter(name == 'tot_eng') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

### Vacancies

```{r}
# Industry vacancies trend
pre19 |>
  filter(name == 'tot_vac') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of vacancies by industry",
    subtitle = "(2000 - 2019)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry vacancies changes
pre19 |>
  filter(name == 'tot_vac') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of vacancies by industry",
    subtitle = "(2000 - 2019)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry vacancies changes
pre19 |>
  filter(name == 'tot_vac') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

```{r}
# Industry vacancies trend
pos19 |>
  filter(name == 'tot_vac') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of vacancies by industry",
    subtitle = "(2019 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry vacancies changes
pos19 |>
  filter(name == 'tot_vac') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of vacancies by industry",
    subtitle = "(2019 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry vacancies changes
pos19 |>
  filter(name == 'tot_vac') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

### Headcounts per establishment

```{r}
# Industry headcounts trend
pre19 |>
  filter(name == 'lab_per_est') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of headcounts per establishment by industry",
    subtitle = "(2000 - 2019)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry headcounts changes
pre19 |>
  filter(name == 'lab_per_est') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of headcounts per establishment by industry",
    subtitle = "(2000 - 2019)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry headcounts changes
pre19 |>
  filter(name == 'lab_per_est') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

```{r}
# Industry headcounts trend
pos19 |>
  filter(name == 'lab_per_est') |>
  ggplot(aes(x=Date,y=value,colour=ind)) +
  geom_line() +
  labs(
    title = "Number of headcounts per establishment by industry",
    subtitle = "(2019 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "No of persons",
    color = 'Industry'
  ) +
  theme_classic()

# Industry headcounts changes
pos19 |>
  filter(name == 'lab_per_est') |>
  group_by(ind) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "% change in number of headcounts per establishment by industry",
    subtitle = "(2019 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry headcounts changes
pos19 |>
  filter(name == 'lab_per_est') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) / first (value) - 1) * 100
  ) |>
  arrange(-change) |>
  print()
```

## Labour market share

```{r}
ind_shr <- ind_lab |>
  group_by(Date) |>
  mutate(
    est_shr = tot_est/sum(tot_est)*100,
    eng_shr = tot_eng/sum(tot_eng)*100,
    vac_shr = tot_vac/sum(tot_vac)*100,
    lab_shr = tot_lab/sum(tot_lab)*100
  ) |>
  select(
    Date,ind,ind_des,est_shr,eng_shr,vac_shr,lab_shr
  )

ind_shr_long <- ind_shr |>
  pivot_longer(-c(Date,ind,ind_des))

head(ind_shr)
head(ind_shr_long)
```

### Establishment

```{r}
# Industry establishment share
ind_shr_long |>
  filter(name == 'est_shr') |>
  ggplot(aes(x=Date,y=value,fill=ind)) +
  geom_area(color='white',linewidth=0.5) +
  labs(
    title = "Share of establishments by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "%",
    color = 'Industry'
  ) +
  theme_classic()

# Industry establishment changes
ind_shr_long |>
  filter(name == 'est_shr') |>
  group_by(ind) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "Change in share of establishments by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry establishment changes
ind_shr_long |>
  filter(name == 'est_shr') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  arrange(-change) |>
  print()
```

### Engaged persons

```{r}
# Industry engagement share
ind_shr_long |>
  filter(name == 'eng_shr') |>
  ggplot(aes(x=Date,y=value,fill=ind)) +
  geom_area(color='white',linewidth=0.5) +
  labs(
    title = "Share of engaged persons by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "%",
    color = 'Industry'
  ) +
  theme_classic()

# Industry engagement changes
ind_shr_long |>
  filter(name == 'eng_shr') |>
  group_by(ind) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "Change in share of engaged persons by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry engagement changes
ind_shr_long |>
  filter(name == 'eng_shr') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  arrange(-change) |>
  print()
```

### Vacancies

```{r}
# Industry vacancies share
ind_shr_long |>
  filter(name == 'vac_shr') |>
  ggplot(aes(x=Date,y=value,fill=ind)) +
  geom_area(color='white',linewidth=0.5) +
  labs(
    title = "Share of vacancies by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "%",
    color = 'Industry'
  ) +
  theme_classic()

# Industry vacancies changes
ind_shr_long |>
  filter(name == 'vac_shr') |>
  group_by(ind) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "Change in share of vacancies by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry vacancies changes
ind_shr_long |>
  filter(name == 'vac_shr') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  arrange(-change) |>
  print()
```

### Total headcounts (primarily engaged)

```{r}
# Industry vacancies share
ind_shr_long |>
  filter(name == 'lab_shr') |>
  ggplot(aes(x=Date,y=value,fill=ind)) +
  geom_area(color='white',linewidth=0.5) +
  labs(
    title = "Share of headcounts by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Date",
    y = "%",
    color = 'Industry'
  ) +
  theme_classic()

# Industry vacancies changes
ind_shr_long |>
  filter(name == 'lab_shr') |>
  group_by(ind) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  ggplot(aes(x=reorder(ind,change),y=change,fill=ind)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  scale_fill_discrete(guide = 'none') +
  labs(
    title = "Change in share of headcounts by industry",
    subtitle = "(2000 - 2026)",
    caption = "Data from CSD HK",
    x = "Industry",
    y = "%"
  ) +
  theme_classic()

# Industry vacancies changes
ind_shr_long |>
  filter(name == 'lab_shr') |>
  group_by(ind_des) |>
  summarize(
    change = (last(value) - first (value))
  ) |>
  arrange(-change) |>
  print()
```
